Exploring probability distributions for bivariate temporal granularities

### 2019-12-08 main.Rmd
discussion change

@@

abstract: |
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible. As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases. We propose some new tools to explore this type of data, which deconstruct time in many different ways. There are several classes of time deconstructions including linear time granularities, circular time granularities and aperiodic calendar categorizations. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. Circular time granularities accommodate periodicities in time such as hour of the day, and day of the week. Aperiodic calendar categorizations are neither linear nor circular, such as day of the month or public holidays.

@@

   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2[@Wickham2009-pk]. 
bibliography: [bibliography.bib]
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

@@

Exploratory data analysis is iteratively finding and summarizing patterns. With temporal data available at more finer scales, exploring periodicity become overwhelming with so many possible cyclic granularities that could be generated for a context. This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by choosing a list of harmonies that need to be plotted together given a hierarchy or calendar structure and recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.

@@

Some guidelines and best practices to explore without wandering aimlessly can save

@@

The authors would like to thank Prof. Antony Unwin for his immense inpu

@@

<!-- # ```{r write-bib} -->
<!-- # write_bib(c("ggplot2", "tidyverse", "sugrrants", "geofacet", "shiny", "wanderer4melb"), 'packages.bib') -->
<!-- # ``` -->

<!-- library(gravitas) -->
<!-- library(tsibble) -->

<!--  cricket_bowling <- cricket_season %>% -->
<!--    mutate(dot_ball_proxy = if_else(total_runs==0,1,0), -->
<!--           wicket_proxy = if_else(is.na(dismissal_kind),0,1)) %>% -->
<!--   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
<!--    summarise(dot_balls = if_else(sum(dot_ball_proxy)>0, "yes","no"), -->
<!--              wickets = if_else(sum(wicket_proxy)>0, "yes","no"),  -->
<!--             total_runs = sum(total_runs)) %>% -->
<!--             mutate(diff_run_rate = c(0,diff(total_runs))) -->
<!--   -->
<!--   -->
<!--  cricket_tsibble_bow <- cricket_bowling %>%  -->
<!--    ungroup() %>%  -->
<!--    mutate(data_index = row_number()) %>%  -->
<!--    as_tsibble(index = data_index) -->

<!-- cricket_tsibble_bow %>% gran_obs("over", "dot_balls") -->

<!-- # cricket_tsibble_bow %>%  -->
<!-- #   #filter(dot_balls %in% c(0, 1, 2))%>%  -->
<!-- #   prob_plot("dot_balls", -->
<!-- #             "over", -->
<!-- #             hierarchy_model, -->
<!-- #             response = "diff_run_rate", -->
<!-- #             plot_type = "quantile", -->
<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
<!-- #             symmetric = FALSE) -->

<!-- cricket_tsibble_bow %>% gran_obs("over", "wickets") -->

<!-- cricket_tsibble_bow %>%  -->
<!--   #filter(wickets %in% c(0, 1))%>%  -->
<!--   gather("proxies", "yes_no", -c(season, -->
<!--                                      match_id, -->
<!--                                      batting_team, -->
<!--                                      bowling_team, -->
<!--                                      inning,     -->
<!--                                      over, -->
<!--                                      diff_run_rate, -->
<!--                                      data_index)) %>%  -->
<!--   filter(yes_no %in% c("no")) %>%  -->
<!--   prob_plot("proxies", -->
<!--             "over", -->
<!--             hierarchy_model, -->
<!--             response = "diff_run_rate", -->
<!--             plot_type = "boxplot", -->
<!--             symmetric = FALSE, -->
<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->
<!-- ``` -->

<!-- ```{r proxies-yes} -->
<!-- cricket_tsibble_bow %>%  -->
<!--   #filter(wickets %in% c(0, 1))%>%  -->
<!--   gather("proxies", "yes_no", -c(season, -->
<!--                                      match_id, -->
<!--                                      batting_team, -->
<!--                                      bowling_team, -->
<!--                                      inning,     -->
<!--                                      over, -->
<!--                                      diff_run_rate, -->
<!--                                      data_index)) %>%  -->
<!--   filter(yes_no %in% c("yes")) %>%  -->
<!--   prob_plot("proxies", -->
<!--             "over", -->
<!--             hierarchy_model, -->
<!--             response = "diff_run_rate", -->
<!--             plot_type = "lv", -->
<!--             symmetric = FALSE, -->
<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ ggtitle("") -->

<!-- ```{r exwicketplot, echo=FALSE} -->
<!-- #  -->
<!-- # cricket_tsibble %>%  -->
<!-- #   filter(dot_balls %in% c(0, 1, 2)) %>% -->
<!-- #   mutate(lag_dot = lag(dot_balls)) %>%  -->
<!-- #   filter(lag_dot %in% c(0, 1, 2)) %>% -->
<!-- #   prob_plot("lag_dot", -->
<!-- #             "over", -->
<!-- #             hierarchy_model, -->
<!-- #             response = "runs_per_over", -->
<!-- #             plot_type = "quantile", -->
<!-- #             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->

<!-- cricket_tsibble %>%  -->
<!--   filter(wicket %in% c(0, 1)) %>% -->
<!--   mutate(lag_wicket = lag(wicket)) %>%  -->
<!--   filter(lag_wicket %in% c(0, 1)) %>% -->
<!--   prob_plot("lag_wicket", -->
<!--             "over", -->
<!--             hierarchy_model, -->
<!--             response = "runs_per_over", -->
<!--             plot_type = "quantile", -->
<!--             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->

<!-- ```{r violin, fig.cap= "Violin and box plots of run rate across innings of the match and matches won or lost for Mumbai Indians(top) and Chennai Super Kings(bottom).The violins for CSK has long right tails in 2nd innings for cases when they lost or won. MI has long right tail only when they won the match."} -->

<!-- matches_select <- matches %>% select(id, toss_winner, winner, win_by_runs) -->

<!-- cricket_winner <- cricket_tsibble_all %>% -->
<!--   left_join(matches_select, by = c("match_id" = "id")) -->

<!-- MI_winner <- cricket_winner %>% -->
<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
<!--                     season %in% c(2013, 2015)) %>% -->
<!--              mutate(win = if_else(winner == "Mumbai Indians", 1, 0)) -->

<!-- MI_winner %>% -->
<!--   gran_obs("win", "inning", hierarchy_model) -->

<!-- CSK_winner <- cricket_winner %>% -->
<!--              filter(batting_team ==  "Chennai Super Kings"| bowling_team == "Chennai Super Kings", -->
<!--                     season %in% c(2013, 2015)) %>% -->
<!--              mutate(win = if_else(winner == "Chennai Super Kings", 1, 0)) -->

<!-- CSK_winner %>% -->
<!--   gran_obs("win", "inning", hierarchy_model) -->

<!--  MI_winner %>% -->
<!--             prob_plot( "win","inning", -->
<!--                                 hierarchy_model, -->
<!--                                 response = "run_rate", -->
<!--                                 plot_type = "violin", -->
<!--                                 quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
<!--                      overlay = FALSE) + ggtitle("") + -->
<!--     geom_boxplot(width = 0.1, color = "blue") -->

<!--  CSK_winner %>% filter(inning %in% c(1,2)) %>% -->
<!--            prob_plot( "win","inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "run_rate", -->
<!--                                plot_type = "violin", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
<!--                     overlay = FALSE) + ggtitle("") + -->
<!--    geom_boxplot(width = 0.1, color = "blue") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cricket_winner %>% as_tibble() %>% -->
<!--              filter(batting_team ==  "Mumbai Indians" | bowling_team == "Mumbai Indians", -->
<!--                     winner!= "Mumbai Indians", -->
<!--                     win_by_runs!=0, -->
<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->

<!-- cricket_winner %>% as_tibble() %>% -->
<!--              filter(batting_team ==  "Chennai Super Kings", -->
<!--                     winner!= "Chennai Super Kings", -->
<!--                     win_by_runs!=0, -->
<!--                     inning ==2, -->
<!--                     season %in% c(2013, 2015)) %>% select(win_by_runs) %>% summary() -->
<!-- ``` -->

<!-- For MI, cases when they lost, distribution of run rates not very different depending on if they bat first or second. However, cases when they won, distribution of run rate is very skewed when they bat in the 2nd innings implying winning by short margins.  -->

<!-- Figure \ref{fig:violin} shows that for CSK, irrespective of the fact they won or lost, distribution of run rate in case they are batting in the 2nd innings is always skewed on the right side, implying that they always manage to give a good fight to win even if they are losing. Run rate can go up to 6 runs per over depending on the situation. For MI, this is not true. When they won, their run rates in the 2nd innings went up to 6 and led to their win, but stayed within 4 whenever they lost. This implies CSK is more of a fighter than MI when it comes to defending in the 2nd innings. However, we can delve deeper and try to see which overs are the most volatile for CSK and if it is different to that of MI? -->

<!-- In Figure \ref{fig:lv}, we can observe the letter value plot describing the distribution of run rate per over across over of the innings and innings of the match. It can be observed that for both innings, the run rate for Mumbai Indians rise faster across overs of the innings. There is an upward shift of the distribution in the second part of the innings (after 10th over) and more variability (longer and distinct letter values) as well. -->
<!-- CSK has been more uniform in their approach maintaining consistent scores per over. Since CSK's last over in the 2nd innings really get volatile, after a consistent performance in the other overs - this behavior is exhibited as an outlier in Figure \ref{fig:violin}. -->

<!-- whereas MI's scores increase as they proceed with the game notwithstanding if they bat first or second. CSK's run rate distribution is extremely right tailed in the 2nd innings, but not in the first. .   --> -->

<!-- ```{r 2013_15_MI} -->
<!--  cricket_tsibble_all %>%  -->
<!--   filter(season %in% c(2013, 2015), batting_team=="Mumbai Indians") %>%  -->
<!--            prob_plot("inning", "over", -->
<!--                                hierarchy_model, -->
<!--                                response = "run_rate", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
<!--     ) -->
<!--  cricket_tsibble_all %>%  -->
<!--   filter(season %in% c(2013, 2015), batting_team=="Chennai Super Kings") %>%  -->
<!--            prob_plot("inning", "over", -->
<!--                                hierarchy_model, -->
<!--                                response = "run_rate", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  + #scale_y_discrete(breaks= seq(1, 25,1)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="white" )  -->
<!--     ) -->

<!-- r lv, fig.cap = "Letter value plot of run rate across overs of an inning and innings of the match for (top) Mumbai Indians and (bottom) Chennai Super Kings."} -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # MI <- cricket_tsibble_all %>% -->
<!-- #   filter(season %in% c(2010:2015), batting_team=="Mumbai Indians") -->
<!-- #  -->
<!-- #  -->
<!-- # CSK <- cricket_tsibble_all %>% -->
<!-- #   filter(season %in% c(2010:2015), -->
<!-- #          batting_team=="Kolkata Knight Riders", -->
<!-- #          inning %in% c(1,2)) -->
<!-- #  -->
<!-- #  -->
<!-- #  MI %>% -->
<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
<!-- #  -->
<!-- #  CSK %>% -->
<!-- #    gran_obs("inning", "over", hierarchy_model) %>% -->
<!-- #    knitr::kable(format = "markdown", caption = "Cross tabulation of Innings and Overs for Mumbai Indians: 2010-2015") -->
<!-- #  -->
<!-- # MI %>% -->
<!-- #   prob_plot("inning", "over", -->
<!-- #                                hierarchy_model, -->
<!-- #                                response = "run_rate", -->
<!-- #                                plot_type = "lv", -->
<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))  + ggtitle("") -->
<!-- #  -->
<!-- #  -->
<!-- # CSK %>% -->
<!-- #   prob_plot("inning", "over", -->
<!-- #                                hierarchy_model, -->
<!-- #                                response = "run_rate", -->
<!-- #                                plot_type = "lv", -->
<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + ggtitle("") -->
<!-- #  -->
<!-- # cricket_tsibble_all %>% distinct(batting_team) -->

<!-- # ```{r dot, fig.cap="Distribution of run rate across overs of the innings and number of dot balls in an over."} -->
<!-- #  -->
<!-- # cricket_dot <- cricket_season %>%  -->
<!-- #   mutate(wicket = if_else(is.na(dismissal_kind),0,1), -->
<!-- #          dot_balls = if_else(total_runs==0,1, 0) -->
<!-- #          ) %>%  -->
<!-- #   group_by(season, match_id, batting_team, bowling_team,  inning, over) %>% -->
<!-- #   summarise(runs_per_over = sum(total_runs), run_rate = sum(total_runs)/length(total_runs), -->
<!-- #             wicket = sum(wicket), -->
<!-- #             dot_balls = sum(dot_balls))  -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # cricket_tsibble_dot <- cricket_dot %>%  -->
<!-- #   ungroup() %>%  -->
<!-- #   mutate(data_index = row_number()) %>%  -->
<!-- #   as_tsibble(index = data_index) -->
<!-- #  -->
<!-- # #cricket_tsibble_dot %>% gran_obs("over", "fielding_wckts") -->
<!-- #  -->
<!-- # cricket_tsibble_dot  %>%  -->
<!-- #   filter(dot_balls %in% c(1,2,3,4)) %>%  -->
<!-- #   mutate(lag_dot_ball = lag(dot_balls)) -->
<!-- #   prob_plot("over","dot_balls", -->
<!-- #                                hierarchy_model, -->
<!-- #                                response = "run_rate", -->
<!-- #                                plot_type = "quantile", -->
<!-- #                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9), -->
<!-- #            overlay = FALSE)  + ggtitle("") -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->

<!--   cricket_tsibble %>% prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + -->
<!--     theme( # remove the vertical grid lines -->
<!--            panel.grid.major.x = element_blank() , -->
<!--            # explicitly set the horizontal lines (or they will disappear too) -->
<!--            panel.grid.major.y = element_line( size=.1, color="black" )  -->
<!--     )

<!--   cricket_tsibble %>% -->
<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
<!--     prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!--     cricket_tsibble %>% -->
<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
<!--     prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!-- ```{r runs_per-over} -->

<!-- # all season per over runs -->
<!-- cricket_per_over_all <- cricket_all_seas %>%  -->
<!--   group_by(season, match_id, batting_team,inning, over) %>% -->
<!--   summarise(runs_per_over = sum(total_runs))  -->

<!-- # all season per over runs tsibble -->
<!-- cricket_tsibble_all <- cricket_per_over_all %>%  -->
<!--   ungroup() %>%  -->
<!--   mutate(data_index = row_number()) %>%  -->
<!--   as_tsibble(index = data_index) -->

<!--   cricket_tsibble_all %>% prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   ) -->

<!--   cricket_tsibble_all %>% prob_plot("inning_match", "season", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )   -->

<!--   cricket_tsibble_all %>% prob_plot("season", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!--   )  +  -->

<!--   cricket_tsibble %>% -->
<!--     filter(batting_team == "Chennai Super Kings") %>%  -->
<!--     prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!--     cricket_tsibble %>% -->
<!--     filter(batting_team == "Kolkata Knight Riders") %>%  -->
<!--     prob_plot("inning_match", "over_inning", -->
<!--                                hierarchy_model, -->
<!--                                response = "runs_per_over", -->
<!--                                plot_type = "quantile", -->
<!--                                quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!--   ) -->

<!-- {r read_data, echo=FALSE} -->

<!-- cricket_tsibble <- cricketdata %>% -->
<!--   mutate(data_index = row_number()) %>% -->
<!--   as_tsibble(index = data_index) -->

<!-- # cricket_tsibble %>% dynamic_create_gran( -->
<!-- #   "ball_over", -->
<!-- #   hierarchy_model -->
<!-- # ) -->

<!-- #over_inning    ball_over -->
<!-- cricket_tsibble %>% granplot("over_inning", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!-- ) -->

<!-- #inning_match   ball_over -->

<!-- cricket_tsibble %>% granplot("inning_match", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!-- ) -->

<!-- # inning_match   ball_inning -->
<!-- #interesting plot -->
<!-- cricket_tsibble %>% granplot("inning_match", "ball_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
<!--   scale_x_discrete(breaks = seq(1,120,5)) -->

<!-- # ball_over      over_inning -->

<!-- cricket_tsibble %>% granplot("ball_over", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->

<!-- # inning_match   over_inning -->

<!-- cricket_tsibble %>% granplot("inning_match", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- # ball_over      over_match -->

<!-- cricket_tsibble %>% granplot("ball_over", "over_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- # ball_over      inning_match -->

<!-- cricket_tsibble %>% granplot("ball_over", "inning_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->

<!-- cricket_tsibble %>% -->
<!--   granplot( -->
<!--     gran1 = "inning_match", -->
<!--     gran2 = "over_inning", -->
<!--     hierarchy_model, -->
<!--     response = "total_runs", -->
<!--     plot_type = "quantile" -->
<!--   ) -->

<!-- ```{r read_data_all_seas, echo=FALSE} -->

<!-- load("/Users/sgup0008/Documents/paper-gravitas/data/cricket_all_seas.rda") -->

<!-- cricket_tsibble_all <- cricket_all_seas %>% -->
<!--   filter(season %in% c(2012:2015)) %>%  -->
<!--   mutate(data_index = row_number()) %>% -->
<!--   as_tsibble(index = data_index) -->

<!-- # cricket_tsibble %>% dynamic_create_gran( -->
<!-- #   "ball_over", -->
<!-- #   hierarchy_model -->
<!-- # ) -->

<!-- #over_inning    ball_over  -->
<!-- cricket_tsibble_all %>% granplot("over_inning", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->

<!-- ) -->

<!-- #inning_match   ball_over  -->

<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_over", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9) -->
<!-- ) -->

<!-- # inning_match   ball_inning  -->
<!-- #interesting plot -->
<!-- cricket_tsibble_all %>% granplot("inning_match", "ball_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--    plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9))+ -->
<!--   scale_x_discrete(breaks = seq(1,120,5)) -->

<!-- # ball_over      over_inning   -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs", -->
<!--   plot_type = "quantile", -->
<!--   quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) -->

<!-- # inning_match   over_inning   -->

<!-- cricket_tsibble_all %>% granplot("inning_match", "over_inning", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- # ball_over      over_match   -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "over_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) -->

<!-- # ball_over      inning_match   -->

<!-- cricket_tsibble_all %>% granplot("ball_over", "inning_match", -->
<!--   hierarchy_model, -->
<!--   response = "total_runs" -->
<!-- ) + scale_color_brewer(type = "div", palette = "Dark2") -->

<!-- cricket_tsibble_all %>% -->
<!--   granplot( -->
<!--     gran1 = "inning_match", -->
<!--     gran2 = "over_inning", -->
<!--     hierarchy_model, -->
<!--     response = "total_runs", -->
<!--     plot_type = "quantile" -->
<!--   ) -->

>>>>

- name: Antony Unwin
  affiliation: University of Augsburg

@@

abstract: |
 Recent advances in technology greatly facilitates recording and storing data at much finer temporal scales than was previously possible. As the frequency of time-oriented data increases, the number of questions about the observed variable that need to be addressed by visual representation also increases. We propose some new tools to explore this type of data, which deconstruct time in many different ways. There are several classes of time deconstructions including linear and cyclic time granularities. Linear time granularities respect the linear progression of time such as hours, days, weeks and months. Cyclic time granularities can be circular such as hour of the day, and day of the week, quasi-circular such as day of the month and aperiodic such as public or school holidays.

@@

   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in ggplot2 [@Wickham2009-pk]. 
bibliography: [bibliography.bib]
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

@@

Exploratory data analysis is iteratively finding and summarizing patterns. With temporal data available at more finer scales, exploring periodicity can become overwhelming with so many possible cyclic granularities that could be generated for a context. This work provides a framework to systematically explore distribution of an observed variable across two cyclic time granularities by creating any cyclic granularity, choosing a list of harmonies and therby recommending possible distribution plots for effective visualization based on relationship and levels of the cyclic granularities.

@@

@@

The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages. This manuscipt was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.

@@
